---
title: Cincinnati Food Safety Program
author: andre
date: '2020-03-04'
slug: cincinnati-food-safety
categories:
  - Public Health
  - R
tags:
  - Food Safety
subtitle: ''
summary: ''
authors: []
lastmod: '2020-03-04T07:53:58-05:00'
featured: no
projects: []
output:
  blogdown::html_page:
    toc: true
---

Image credit: [**Unsplash**](https://unsplash.com/photos/CpkOjOcXdUY)

* <span style="color: #fd7f6f;">Complaint</span>

* <span style="color: #7eb0d5;">Routine</span>

* <span style="color: #b2e061;">Initial</span>

# Setup

```{r setup, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library("RSocrata")
library(tidyverse)
library(knitr)
library(DBI)
library(plotly)
library(htmlwidgets)
library(htmltools)
library(leaflet)
```


# Getting data

First, you will need to create your database. Click [here](https://andrevanzyl.netlify.com/post/install-postgresql-on-kubuntu-ubuntu-19-10/) for instructions on setting up a PostgreSQL server on Kubuntu 19.10. 

```{r echo=TRUE}
if (Sys.info()["sysname"] != "Linux"){
  con <- dbConnect(odbc::odbc(), Server = "127.0.0.1", Database = "cincinnati", 
    UID = "dre", PWD = "password", Port = 5432, .connection_string = "Driver={PostgreSQL ODBC Driver(ANSI)};", 
    timeout = 10)
  
} else {
  con <- dbConnect(odbc::odbc(), Server = "127.0.0.1", Database = "cincinnati", 
    UID = "dre", PWD = "password", Port = 5432, .connection_string = "Driver={PostgreSQL ANSI};", 
    timeout = 10)
}
```

```{r message=FALSE, warning=FALSE, include=TRUE}

## Install the required package with:
## install.packages("RSocrata")


if (!dbExistsTable(con, "foodsafety")){
  df <- read.socrata(
  "https://data.cincinnati-oh.gov/resource/rg6p-b3h3.json"#,
  #app_token = "YOURAPPTOKENHERE",
  #email     = "user@example.com",
  #password  = "fakepassword"
)
  dbWriteTable(con, "foodsafety", df, row.names=FALSE, append=FALSE)
}
dbExistsTable(con, "foodsafety")
```


# Column Names

```{sql connection=con, output.var = "results" }
SELECT column_name,data_type 
FROM information_schema.columns 
WHERE table_name = 'foodsafety';
```

```{r message=FALSE, warning=FALSE}
results %>% knitr::kable()
```

# A Map

```{sql connection=con, output.var = "results"}
ALTER TABLE foodsafety
  ALTER COLUMN "latitude" TYPE NUMERIC(14, 11) 
    USING ("latitude"::NUMERIC(14,11)),
  ALTER COLUMN "longitude" TYPE NUMERIC(14, 11) 
    USING ("longitude"::NUMERIC(14,11));
```

```{sql connection=con, output.var = "results" }
SELECT  DISTINCT(postal_code), latitude, longitude, insp_type
FROM foodsafety
```

```{r message=FALSE, warning=FALSE}
pal <- colorFactor(c("#fd7f6f", "#7eb0d5", "#b2e061"), domain = c("COMPLAINT", "ROUTINE", "INITIAL"))

leafMap <- leaflet(results) %>% addTiles() %>%
  addCircleMarkers(
    lng=~longitude, lat=~latitude,
    color = ~pal(insp_type),
    stroke = FALSE, fillOpacity = 0.5,
    clusterOptions = markerClusterOptions()
  ) %>% addProviderTiles(providers$Esri.NatGeoWorldMap)
```

```{r eval=FALSE, include=FALSE, echo=TRUE}
leafMap
```


```{r}
currentWD <- getwd()
dir.create("static/leaflet", showWarnings = FALSE)
setwd("static/leaflet")
saveWidget(leafMap, "leafMap.html")
setwd(currentWD)
```


<iframe seamless src="/leaflet/leafMap.html" width="100%" height="500"></iframe>


# Using dplyr

```{r}
foodsafety <- tbl(con, "foodsafety")

foodsafety %>% group_by(insp_type) %>% summarise(counts = n()) %>% show_query()
```

```{sql connection=con, output.var = "results" }
SELECT insp_type, COUNT(*) AS counts
FROM foodsafety
GROUP BY insp_type
```

```{r message=FALSE, warning=FALSE}
results %>% knitr::kable()
```




```{sql connection=con, output.var = "results" }
SELECT  DISTINCT(postal_code), latitude, longitude, insp_type
FROM foodsafety
WHERE insp_type = 'COMPLAINT'
```

```{r eval=FALSE, include=FALSE}
library(leaflet)
#pal <- colorFactor(c("#fd7f6f", "#7eb0d5", "#b2e061"), domain = c("COMPLAINT", "ROUTINE", "INITIAL"))

leaflet(results) %>% addTiles() %>%
  addCircleMarkers(
    lng=~longitude, lat=~latitude,
    color = ~pal(insp_type),
    stroke = FALSE, fillOpacity = 0.5,
    clusterOptions = markerClusterOptions()
  ) %>% addProviderTiles(providers$Esri.NatGeoWorldMap)
```

# Sample of Data

```{sql connection=con, output.var = "results"}
SELECT *
FROM foodsafety
LIMIT 5
```


```{r message=FALSE, warning=FALSE}
results %>% kable()
```

# Duplicate rows?


```{sql connection=con, output.var = "results" }
SELECT recordnum_license,	license_no,	business_name,	address,	city,	state,	postal_code,	latitude,	longitude,	phone_number,	license_status,	recordnum_insp,	insp_type,	insp_subtype,	action_date,	action_sequence,	action_status,	code,	violation_description,	violation_comments,	violation_key,	last_table_update,	neighborhood, COUNT(*) AS counts
FROM foodsafety
GROUP BY recordnum_license,	license_no,	business_name,	address,	city,	state,	postal_code,	latitude,	longitude,	phone_number,	license_status,	recordnum_insp,	insp_type,	insp_subtype,	action_date,	action_sequence,	action_status,	code,	violation_description,	violation_comments,	violation_key,	last_table_update,	neighborhood
--HAVING
--    COUNT(*)>1
ORDER BY counts DESC
LIMIT 3;
```

```{r message=FALSE, warning=FALSE}
results %>% knitr::kable()
```

# Neighborhood


```{sql connection=con, output.var = "results" }
SELECT neighborhood, COUNT(*) AS counts
FROM foodsafety
GROUP BY neighborhood
ORDER BY counts DESC
LIMIT 20
```

```{r message=FALSE, warning=FALSE}
results %>% knitr::kable()
```


# Business name


```{sql connection=con, output.var = "results" }
SELECT business_name, COUNT(*) AS counts
FROM foodsafety
GROUP BY business_name
ORDER BY counts DESC
LIMIT 10
```

```{r message=FALSE, warning=FALSE}
results %>% knitr::kable()
```


```{sql connection=con, output.var = "results" }
SELECT last_table_update, business_name, violation_comments
FROM foodsafety 
WHERE business_name
LIKE '"BRIDGES%'
ORDER BY last_table_update DESC
LIMIT 10
;
```

```{r message=FALSE, warning=FALSE}
results %>% knitr::kable()
```


# Action dates


```{sql connection=con, output.var = "results" }
SELECT 
  * 
FROM 
  foodsafety 
  -- Return rows where demand_loss_mw is not missing or unknown   
WHERE 
  action_date IS NULL;
```


```{r message=FALSE, warning=FALSE}
results %>% knitr::kable()
```


```{sql connection=con, output.var = "results" }
SELECT DISTINCT action_date, insp_subtype, business_name
FROM foodsafety
ORDER BY business_name, action_date
LIMIT 10
```


```{r message=FALSE, warning=FALSE}
results %>% knitr::kable()
```


```{sql connection=con, output.var = "results" }
SELECT DISTINCT action_date, insp_subtype, business_name
FROM foodsafety
ORDER BY business_name, action_date DESC
LIMIT 10
```


```{r message=FALSE, warning=FALSE}
results %>% knitr::kable()
```


```{sql connection=con, output.var = "results" }
SELECT DISTINCT action_date, insp_subtype, business_name
FROM foodsafety
WHERE action_date >= '2020-03-01 00:00:00'::timestamp
ORDER BY action_date DESC
```


```{r message=FALSE, warning=FALSE}
results %>% knitr::kable()
```



```{r eval=FALSE, include=FALSE}
foodsafety %>% 
  select(action_date, insp_type) %>%
  filter(!is.na(action_date), insp_type == "COMPLAINT") %>%
  mutate(action_month = month(action_date, label = FALSE)) %>% 
  group_by(action_month) %>% 
  arrange(action_month) %>% 
  summarise(complaint_actions = n()) %>% show_query()
```


```{sql connection=con, output.var = "results" }
SELECT "action_month", COUNT(*) AS "complaint_actions"
FROM (SELECT *
FROM (SELECT "action_date", "insp_type", EXTRACT(MONTH FROM "action_date") AS "action_month"
FROM (SELECT *
FROM (SELECT "action_date", "insp_type"
FROM "foodsafety") "dbplyr_117"
WHERE ((NOT((("action_date") IS NULL))) AND ("insp_type" = 'COMPLAINT'))) "dbplyr_118") "dbplyr_119"
ORDER BY "action_month") "table"
GROUP BY "action_month"
```

```{r message=FALSE, warning=FALSE}
results %>% knitr::kable()
```

```{r message=FALSE, warning=FALSE}
fig <- plot_ly(x = ~results$action_month, y = ~results$complaint_actions, mode = 'lines', text = paste("seconds from now in"))

root <- getwd()
setwd("static/plots/fs/complaint_actions/")
htmlwidgets::saveWidget(as_widget(fig), "complaint_actions.html")
setwd(root)
```

```{r eval=FALSE, include=FALSE}
fig
```


<iframe seamless src="/plots/fs/complaint_actions/complaint_actions.html" width="100%" height="500"></iframe>


# Inspection Types 

```{sql connection=con, output.var = "results" }
SELECT DISTINCT insp_type
FROM foodsafety
```


```{r message=FALSE, warning=FALSE}
results %>% knitr::kable()
```



# Cities 

```{sql connection=con, output.var = "results" }
SELECT DISTINCT city AS Cities
FROM foodsafety
```


```{r message=FALSE, warning=FALSE}
results %>% knitr::kable()
```

# Postal Codes 

```{sql connection=con, output.var = "results"}
SELECT DISTINCT postal_code AS postal_codes, city
FROM foodsafety
LIMIT 10
```

```{r message=FALSE, warning=FALSE}
results %>% knitr::kable()
```

```{sql connection=con, output.var = "results"}
SELECT 
  neighborhood,
  business_name,
  last_table_update,
  violation_description,
  postal_code
FROM 
  foodsafety
WHERE 
  ((
    business_name LIKE '"BRIDGE%' 
    AND postal_code = '45223'
  AND last_table_update >= '2019-10-01'::timestamp
  ) 
  OR (postal_code = '45220' AND
  business_name LIKE '"BR%' AND
  last_table_update >= '2019-10-01'::timestamp))
ORDER BY 
  neighborhood,
  business_name,
  violation_description;
```

```{r message=FALSE, warning=FALSE}
results %>% knitr::kable()
```



# Complaint Discriptions

```{sql connection=con, output.var = "results"}
SELECT DISTINCT violation_description AS complaint_description
FROM foodsafety
WHERE insp_type = 'COMPLAINT'
LIMIT 5
```

```{r message=FALSE, warning=FALSE}

if (Sys.info()["sysname"] != "Linux"){
  results
  
} else {
  results %>% kable()
}

```


# Build Site

```{r eval=FALSE, include=FALSE}
if (exists("con")) {
  library(DBI)
  dbDisconnect(con)
}

con.State
isPostgresqlIdCurrent(con)
dbCanConnect(con)

  setwd("/home/dre/Documents/blog/")
  blogdown::build_site()
  blogdown::serve_site()
  servr::daemon_stop(1)
```

The html that was saved with the saveWidget function can be called with the iframe html tag.

* <span style="color: #fd7f6f;">Complaint</span>
* <span style="color: #7eb0d5;">Routine</span>
* <span style="color: #b2e061;">Initial</span>

# References

[Food Safety and Inspections](https://www.cincinnati-oh.gov/health/cincinnati-health-department-programs/food-safety-and-inspections-program/)

[Cincinnati Food Safety Program](https://data.cincinnati-oh.gov/Thriving-Neighborhoods/Cincinnati-Food-Safety-Program/rg6p-b3h3)

[SODA: Cincinnati Food Safety Program](https://dev.socrata.com/foundry/data.cincinnati-oh.gov/rg6p-b3h3)

[Connect to a Database](https://db.rstudio.com/getting-started/connect-to-database/)

[Databases using R](https://db.rstudio.com/)

[How to get a list column names and datatype of a table in PostgreSQL?](https://stackoverflow.com/questions/20194806/how-to-get-a-list-column-names-and-datatype-of-a-table-in-postgresql)

[Open A Socrata Dataset In Microsoft Power BI](https://support.socrata.com/hc/en-us/articles/115011432428)

[SQL Tutorial](https://www.w3schools.com/sql/default.asp)

[Leaflet for R](https://rstudio.github.io/leaflet/)

[Postgres columns of latitude and longitude from varchar to numeric](https://dba.stackexchange.com/questions/78580/postgres-columns-of-latitude-and-longitude-from-varchar-to-numeric)

[Using Leaflet and htmlwidgets in a Hugo-generated page](https://waterdata.usgs.gov/blog/leaflet/)

[PostgreSQL- Filter a date range](https://stackoverflow.com/questions/15817871/postgresql-filter-a-date-range)

[Databases using dplyr](https://db.rstudio.com/dplyr/)

[Time Series in R](https://plot.ly/r/time-series/)